---
title: "stat 425 hw2"
author: "Cheng Chen"
date: "2023-07-07"
output: html_document
---

Problem 1
```{r}
library(faraway)
data(sat)
model1 <- lm(total ~ expend + ratio + salary, data = sat)
```

(a)
```{r}
summary(model1)$coefficients
```
p-value of "salary" is 6.666771e-02 > 0.05, we do not reject H0: βsalary = 0


(b)
```{r}
anova(model1)
```
reject null hypothesis, because p-value of "expend" is less than 0.05
"ratio" has an effect onmthe response


(c)
```{r}
model2 <- lm(total ~ expend + ratio + salary + takers, data = sat)
summary(model2)$coefficients
anova(model1, model2)
```
p-value is 2.607e-16 < 0.05, we reject null hypothesis
the p-value from the t-test matches the p-value from the F-test in the ANOVA table
so the F -test here is equivalent to the t-test


Problem 2
```{r}
library(faraway)
data(prostate)
model1 <- lm(lpsa ~ ., data = prostate)
```

(a)
```{r}
confint(model1, "age", level = 0.90)
confint(model1, "age", level = 0.95)
```

(b)
```{r}
reduced_model <- step(model1, direction = "backward", alpha = 0.05)
```
```{r}
anova(reduced_model, model1)
```
p-value is 0.4421 > 0.05, original model is preferred over reduced model.


(c)
```{r}
library(ellipse)
```

```{r}
cov_matrix <- vcov(model1)
cov_age_lbph <- cov_matrix[c("age", "lbph"), c("age", "lbph")]
confidence_ellipse <- ellipse(cov_age_lbph, centre = coef(model1)[c("age", "lbph")], level = 0.95)
plot(confidence_ellipse, type = "l", xlab = "age", ylab = "lbph")
points(0, 0, pch = 19)
```
The confidence region includes the origin, so we do not reject the null hypothesis.


(d)
```{r}
n <- nrow(prostate)
t_obs <- summary(model1)$coef[4,3]
n_perm <- 10000
t_perm <- numeric(n_perm)
for (i in 1:n_perm) {
  age_perm <- sample(prostate$age, n, replace = FALSE)
  data_perm <- prostate
  data_perm$age <- age_perm
  model_perm <- lm(lpsa ~ ., data = data_perm)
  t_perm[i] <- summary(model_perm)$coef[4,3]
}
p_value <- sum(abs(t_perm) >= abs(t_obs))/n_perm
p_value
```


Problem 3
```{r}
data(prostate)
model <- lm(lpsa ~ ., data = prostate)
```

(a)
```{r}
new_patient <- data.frame(
  lcavol = 1.44692, lweight = 3.62301, age = 65, lbph = 0.3001,
  svi = 0, lcp = -0.79851, gleason = 7, pgg45 = 15
)
prediction1 <- predict(model, newdata = new_patient, interval = "confidence", level = 0.95)
prediction1
```

(b)
```{r}
new_patient$age <- 20
prediction2 <- predict(model, newdata = new_patient, interval = "confidence", level = 0.95)
prediction2
```

(c)
```{r}
summary1 <- summary(model)
significant_vars <- rownames(summary1$coefficients)[summary1$coefficients[,4] < 0.05]
reduced_model <- lm(lpsa ~ ., data = prostate[, c(significant_vars, "lpsa")])
new_patient <- new_patient[, significant_vars]
prediction3 <- predict(reduced_model, new_patient, interval = "confidence", level = 0.95)
print(prediction3)
```
The CI is narrower, the model with only the significant predictors provides a more precise prediction, so the reduced model is preferred.


Problem 4
```{r}
library(car)
```









